Limited evidence for widespread heterogeneity in psychology

### 2018-06-06 writing/draft-heterogeneity.rmd
Updated paper draft

---
title: "Reviewing evidence for heterogeneity in psychology"
author: "Anton Ohlsson Collentine"
header-includes:
   - \setlength\parindent{24pt}
output: pdf_document
---

@@

#Add rs to the data-frame
het <- dat %>% 
  select(rs, Site, effect, effect_type) %>% 
  group_by(rs, effect) %>% 
  summarize(effect_type = unique(effect_type),
            k = n_distinct(Site)) %>% 
  left_join(res) %>% 
  mutate(effect_type = recode(effect_type, 
                              d = "SMD.",
                              'Raw mean difference' = "MD",
                              'Risk difference' = "RD")) %>% 
  select(rs, effect,k , everything()) %>% 
  arrange(desc(s_I2), desc(s_ci.ub)) #sort by I2 and upper CI bound
```
Empirical research typically proceeds in two stages. First, belief in the existence of an effect is established. Second, its generalizability is examined by exploring its boundary conditions. In the first stage, inferential statistics are used to minimize the risk that a discovery is due to sampling error. In the second stage, we ask to what extent the effect depends on a particular choice of four contextual factors; the 1) sample population, 2) settings, 3) treatment variables and 4) measurement variables [ref: Campbell & Stanley]. This extent is often explored through replications of the original study with some variation, and once sufficient studies have accumulated through meta-analysis. In meta-analysis, the heterogeneity of an effect size (henceforth referred to as heterogeneity) is a measure of an effect's susceptibility to changes in these four factors. An effect strongly dependent on one or more of the four factors, unless controlled for, should exhibit high heterogeneity. Heterogeneity is often measured by the I2 index [ref: Higgins & Higgins], which allows comparison of estimates across meta-analyses and has an intuitive interpretation. It can be interpreted as the percentage of variability in effect sizes that is due to heterogeneity amongst the true effects (that is, dependency on contextual factors) rather than sampling variance, and ranges from 0-100%.

Heterogeneity is of concern for several reasons. First and foremost, unaccounted for heterogeneity can have unacceptable consequences. This may be most readily evident for medicine, where in the case of heterogeneity an intervention that is successful for some may have direct negative health consequences for others, but the adoption of psychological research for business, law enforcement, and policymaking in general (e.g., the widespread use of 'nudges'; https://www.knowablemagazine.org/article/society/2018/nudging-grows-and-now-has-government-job) means its impact can be widespread and the consequences of heterogeneity should be no less a concern. Second, unaccounted for heterogeneity is an indication of incomplete theory, since it indicates that a theory is unable to predict all contextual factors of importance to its claims. While heterogeneity reveals that a theory has flaws, by itself it does not damn a theory, and can be seen as an opportunity for theoretical advancement [ref: Tacket et al]. Third, heterogeneity can create controversy in the interpretation of replication results. The lack of a clear line for the degree of heterogeneity (susceptibility to contextual factors) that would make an effect uninteresting often results in heated discussions between original and replication authors. In response to this common polemic [ref Simons et al] recently proposed that authors include a statement to explicitly delineate the level of heterogeneity that would make an effect lose its value. Fourth, meta-analytic techniques that attempt to correct for publication bias in their estimate tend to fail in the presence of heterogeneity [refs: Inzlicht, McShane]. This is problematic considering the widespread publication bias [refs] and its biasing impact on meta-analytic estimates. Finally, heterogeneity affects the interpretation of meta-analytic estimates as either _the_ true effect size (under homogeneity) or the average of the true effect sizes (under heterogeneity)

It is a common belief that heterogeneity is the norm in psychology. For example, several authors [ref: Tacket et al, McShane et al] argue that if we would expect heterogeneity to be absent or minimal anywhere, it would be in pre-registered multi-lab studies with a common protocol, and that heterogeneity has been reported even under such circumstances [e.g., ref: Many Labs 1] is an indication of widespread heterogeneity in psychology. In addition, recent large scale reviews of meta-analyses in psychology [ref: Stanley; van Erp et al] report that much variability in effect sizes can be attributed to heterogeneity, with median estimates of 74% and 70.62% respectively [ref: Stanley]. In comparison, the median heterogeneity estimate in medicine is considerably lower; 21% amongst Cochrane meta-analyses [ref: Ioannidis  "uncertainty in heterogeneity"]. However, in our experience psychologist tend to include widely varying studies in their meta-analyses, sometimes varying on all four contextual factors (or more), whereas this is less common in medicine. This is perhaps unsurprising considering the lack of delineating theory in psychology [ref: Simons et al]. In support of this notion, Ioannidis reports that, even after excluding meta-analyses of 3 or fewer studies, the median number of studies included in Cochrane meta-analyses is only 7, whereas in the dataset of [ref: van Erp et al] containing 747 meta-analyses in psychology, the median number of studies (effects) per meta-analysis is 12.

Heterogeneity (I2) estimates typically contain considerable uncertainty. The I2 index is a transformation of the Q-test [ref: Cochran] by taking $[Q - (k - 1)] / Q \times 100$%, where k is the number of studies included in the meta-analysis and a negative I2 is set to zero. However, both the Q-test and I2 have low power [ref: Jackson; Huedo-Medina]. This complicates the discussion of heterogeneity, because while I2 always provides an estimate of heterogeneity, this estimate is typically surrounded by large uncertainty [ref: Ioannidis  "uncertainty in heterogeneity"]. For example, Ioannidis reports that in a large set of Cochrane meta-analyses all studies with I2 point estimates of 0% had upper 95% confidence intervals that exceeded 33%. In addition, under homogeneity Q, and hence I2, has a central chisquare distribution [ref: von Hippel], a distribution that is right-skewed with more than 40% of the distribution falling above the expected value in almost all cases (for all k > 4). In other words, even in the absence of heterogeneity, a meta-analysis of 5 or more studies will have an I2 estimate above zero in more than 40% of cases. Point estimates (expected values) of heterogeneity may thus be congruent with a wide range of true heterogeneity levels. It is unfortunately common to not include confidence intervals in the reporting of I2 [e.g., Ml1 and RRRX].

Larger effect sizes can be expected to lead to more heterogeneity. Intuitively, it makes sense to think that if there is no meta-analytic effect there is nothing to moderate (i.e., no heterogeneity). However, a null or near null effect size estimate may arise from failure to consider contextual factors ('hidden moderators', [ref: van Bavel]) and does not by itself imply the absence of heterogeneity. This is especially true when the uncertainty in I2 estimates (see above) is taken into account. A large effect size on the other hand, can be expected to lead to more heterogeneity. To illustrate, consider a meta-analysis of say, the correlation between neuroticism and procrastination. Each included study would need to measure the two variables somehow, hopefully the same way across studies. However, because of individual differences, measurement is not equally exact for everyone, and between studies the measurement reliabilities will differ either due to sampling variance (that the sample happens to be more or less homogenous) or to differences in contextual factors (e.g., sampling population, method of measurement). This means that even if the underlying effect size is the same, the observed correlation between the two variables will differ between studies. If measurement reliabilities stays the same, these differences will increase with the underlying effect, resulting in more variability being ascribed to heterogeneity. More formally, an observed correlation $r_{xy}$ is equivalent to the true correlation $\rho_{xy}$ multiplied by the square root of the measurement reliabilities for X ($R_{xx'}$) and Y ($R_{yy'}$): $r_{xy} = \rho_{xy} \times \sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$. As such, if $\sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$ stays constant while $\rho_{xy}$ increases, the observed differences between studies will increase (see Table 2).

@@

```
We explore heterogeneity across 10 pre-registered multi-lab replication projects in psychology (Table 2). In light of the uncertainty surrounding I2-estimates we review what can be learnt about heterogeneity in psychology from these projects, as well as explore the relationship between effect size and heterogeneity. These 10 projects examine a total of 37 effects and arguably represent the best, least biased, meta-analytic data currently available in psychology. All included studies vary on a minimal number of contextual factors [as recognized by e.g., Tacket et al]: only sample population and settings. Nonetheless, if an effect is sensitive to changes in these factors we may still expect to find some heterogeneity. On the other hand, heterogeneity estimates will only inform us of susceptibility to changes in these two factors, with no information on susceptibility to changes in other contextual factors.

@@

summary.table %>% knitr::kable("latex", booktabs = T, digits = 0) %>% 
  kable_styling() %>% 
  column_spec(2, width = "25em") %>% #specify width of paper column
  footnote(threeparttable = TRUE,
           general = "For studies with several effects the number of participants is the average across effects, rounded to the closest whole number. Participant numbers are those used for primary analyses by original authors (i.e., after exclusions). RS = Replication Study, k = no. primary studies, ML = Many Labs, RRR = Registered Replication Report.")
```

@@

All code and data for this project is available on the Open Science Framework (OSF) at osf.io/XXXX. We refer to relevant files on OSF using brackets and links in the sections below, that is (osf.io/XXXX/). We ran all analyses using R version X [ref: R citation].

We downloaded and collated summary data from 10 pre-registered multi-lab replication projects in psychology (Table 2). Data from all 10 projects were publicly available on the Open Science Framework (osf.io). Although some projects (e.g. RRR4) reported results from several outcome variables, we only included primary outcome variables as stated in accompanying publications, resulting in a total of 37 effects. For each effect we extracted (osf.io/XXXX/) summary data (e.g., means and standard deviations) at the level of the lab as specified by the original authors for their primary analysis (i.e., typically after exclusions). We extracted information on the country of each lab, whether participants were physically in the lab for the study, total number of participants per lab, type of effect size and additional information related to each effect (see codebook; osf.io/XXXX/). Extracted data were in a variety of formats: Excel (Many labs 1, RRR1 & RRR2), CSV (Many labs 3, RRR3, RRR4, RRR5, RRR6) and as PDF tables (RRR7, RRR8). In two cases (RRR5 and RRR6) it was necessary to download the raw data to extract their summary data. Although a particular lab may have participated in several projects, the lab indicator was typically not the same across projects. Even so, we kept the original lab indicators to facilitate comparing observations in our dataset with the original datasets. The summary data for all effects were finally collated into one dataset for analysis (osf.io/XXXX/).

To examine heterogeneity across effects we computed meta-analytic estimates for all 37 effects in our dataset (Table 3). We ran each analysis using the specifications of the replication project authors, with the addition of the Knapp & Hartung adjustment [ref] if it was not already used, since we were primarily interested in heterogeneity estimates (osf.io/XXXX/). All studies used random-effects models, though with a variety of outcome variable types: Correlations, standardized mean differences, raw mean differences, odds ratios and risk differences. For each effect we estimated I2 and its 95% confidence interval.

To examine the relationship between effect size and I2 we converted all effect sizes to a common metric (osf.io/XXXX/). We first converted all effects into correlations (r) and used the R-package metafor [ref] to estimate I2 and meta-analytic effects expressed as Fisher's z. In doing so, we fit random-effects models with the Knapp and Hartung adjustment and metafor's default estimator of Restricted Maximum Likelihood. Fisher's z is a normalizing correlation transformation that ranges from negative to positive infinity, though except for extreme correlations it stays close to the -1 to 1 range [ref]. For mean differences we calculated the pooled standard deviation, [ref Cooper & Hedges, 1994, p. 226), standardized the effect size and converted it to a correlation (p. 234) with a correction factor for unequal sample sizes (p. 234). For risk differences and odds ratios we first added 1/2 to a cell if it was empty to avoid dividing by zero [ref?], next calculated the logarithmic odds ratio (p. 266), converted this to Cohen's d (p. 232) and then to a correlation. All formulas are presented in Appendix A.

@@

##Table 3.
```{r heterogeneity-table}
library(kableExtra)
het %>% knitr::kable("latex", align = "c", booktabs = TRUE,digits = 2, col.names = c("RS", "Effect", "k", "Effect type", "Size", "I2(%)", "ci.lb", "ci.ub")) %>% 
  kable_styling("scale_down") %>%
  footnote(threeparttable = TRUE, escape = FALSE,
           general = "We meta-analyzed effects using the same outcome measures that were used by the original authors. Standardized mean difference is Hedge's g. The following effects are odds ratios transformed into standardized mean differences: 'Allowed vs. forbidden', 'Gain vs. loss framing', 'Norm of reciprocity', 'Low vs. high category scales'. Effects were estimated in metafor using REML with the 'knha' correction. RS = Replication Study, k = no. primary studies, Size = Point estimates of effect sizes, ci.lb = lower bound of I2 95\\\\% confidence interval, ci.ub = upper bound of I2 95\\\\% confidence interval. SMD = Standardized Mean difference (Hedge's g), MD = Mean Difference, RD = Risk Difference, r = correlation.")

@@

What can be learnt from the null I2 point estimates? As can be seen from table 3, 13/37 (35%) I2 point estimates are zero. If there was no heterogeneity across all effects we would expect approximately 50-55% of point estimates to be zero (based on the chisquare distribution and k across studies), that is around 20/37. This supports that there is some heterogeneity across the studied effects. Looking at the confidence intervals of the null estimates we see that in all cases the lower bounds of their confidence intervals are zero, whereas the upper bound varies from 13.75% (Stroop effect) up to 70.62% (Grammar on intention attribution). The three with the highest upper bounds (70.62%, 54.49%, 53.18%) belong to the two replication studies with the lowest k, possibly indicating a lack of power. [ref: Higgins et al] tentatively defined 25% as low heterogeneity, 50% as moderate and 70% as high. Of the upper confidence bounds of the null effects 8/13 lie closest to Higgin's low heterogeneity, 4/13 to moderate, and 1/13 to high. Thus of the null estimates eight are likely low in heterogeneity, whereas there is considerable uncertainty for the remaining five.

What can be learnt from the non-null I2 point estimates? The non-null point estimates of heterogeneity vary from 0.01% to 91.29%. Of these, 16/24 have point estimates that are closest to 'low' heterogeneity, 4/24 to medium and 4/24 to high. Seventeen (71%) have lower confidence bounds of zero heterogeneity and all have upper confidence bounds approaching medium/large heterogeneity. In the majority of non-null I2 point estimates there is thus large uncertainty surrounding heterogeneity levels.

What can be learnt from the significant I2 estimates? Seven out of the 37 effects (19%) have confidence intervals whose lower bounds exclude zero. The point estimates of these effects all approach at least medium heterogeneity, varying from 40.05% to 91.29%. Three of the effects have lower confidence bounds closest to low heterogeneity, and the other four to either medium or high heterogeneity. The upper confidence bounds of all seven are consistent with high heterogeneity. That is, we can be confident that heterogeneity is medium - large for four effects, whereas the remaining three estimates are consistent with anything from low to high heterogeneity.

What have we learnt from the 10 pre-registered multi-lab replication projects in Table 3 about heterogeneity in psychology? Because the included effects are neither a representative nor random sample of effects we cannot make broad claims about heterogeneity in psychology based on these data. Nonetheless, we can with some confidence claim that amongst the examined effects there is heterogeneity, or we would have expected to find a higher proportion of null I2 estimates. Examining the 95% confidence intervals, we can be confident that 8/37 (22%) have low heterogeneity (as defined by [ref: Higgins]), 4/37 (11%) have moderate to high heterogeneity whereas for the remaining 25 effects (68%) there is considerable uncertainty, with estimates consistent with low-moderate or low-high heterogeneity. Because the replication projects only varied on sampling population and settings, we know nothing about the effects' susceptibility to changes in treatment or measurement variables.

Larger effect sizes appear to lead to higher heterogeneity estimates (Figure 1). In Table 3, the four effects with highest I2 estimates also have the largest effect sizes. As expected, heterogeneity tends to increase with (absolute) effect size, with effect size sufficient to explain 61% of variation in I2 estimates according to a simple regression model. The I2 index is truncated at zero, however, the closely related but unbounded heterogeneity estimate H2 provides a similar result (R2 = .53).

@@

<!-- General thought: Ml1 compares heterogeneity from the perspective of between lab vs. between effects and concludes that heterogeneity depends mostly on the effect. However, it seems unsurprising that a certain lab (when not perversely motivated) does not systematically bias effects, but rather that a location might have differential effects on different studied effects. -->

<!-- This might mean that we have missed an important moderator, or that there is a lot of small moderators missed (thinking about it, this seems to differ from random fluctuations in that they are still systematic in some sense, on the other hand, whether they cancel out or not is in itself due to "chance". If an effect (but perhaps not others) is moderated by what time of the day the study is run this can lead to heterogeneity, but say there is also a time-independent moderator such as whether the room is warm -> then we get heterogeneity even if two labs measure at the same time, at what point do these things become "noise" rather than systematic bias? Because we have so few labs it is difficult to separate these?) -->

>>>>

---
title: "Limited evidence for widespread heterogeneity in psychology"
author: "Anton Ohlsson Collentine"
header-includes:
   - \setlength\parindent{24pt}
   - \usepackage{setspace}
   - \doublespacing
output: pdf_document
---

@@

#Add rs to the data-frame
het <- dat %>% 
  select(rs, Site, effect, effect_type) %>% 
  group_by(rs, effect) %>% 
  summarize(effect_type = unique(effect_type),
            k = n_distinct(Site)) %>% 
  left_join(res) %>% 
  mutate(effect_type = recode(effect_type, 
                              d = "SMD.",
                              'Raw mean difference' = "MD",
                              'Risk difference' = "RD")) %>% 
  select(rs, effect,k , everything()) %>% 
  arrange(desc(s_I2),  desc(s_ci.ub)) #sort by I2 and upper CI bound
```
Empirical research typically proceeds in two stages. First, belief in the existence of an effect is established. Second, its generalizability is examined by exploring its boundary conditions. In the first stage, we use inferential statistics  to minimize the risk that a discovery is due to sampling error.  In the second stage, we ask to what extent the effect depends on a particular choice of four contextual factors; the 1) sample population, 2) settings, 3) treatment variables and 4) measurement variables [ref: Campbell & Stanley]. This extent is often explored through replications of the original study with some variation, and once sufficient studies have accumulated through meta-analysis. In meta-analysis, the heterogeneity of an effect size (henceforth referred to as heterogeneity) is a measure of an effect's susceptibility to changes in these four factors. An effect strongly dependent on one or more of the four factors, unless controlled for, should exhibit high heterogeneity. In this paper we examine what degree of heterogeneity can be expected in replication studies in psychology and explore a proposed relationship with effect size.

Heterogeneity is of concern for several reasons. First and foremost, unaccounted for heterogeneity can have unacceptable consequences. This may be most readily evident for medicine, where in the case of heterogeneity an intervention that is successful for some may have direct negative health consequences for others. However, the adoption of psychological research for business, law enforcement, and policymaking in general (e.g., the widespread use of 'nudges'; https://www.knowablemagazine.org/article/society/2018/nudging-grows-and-now-has-government-job) means its impact can be widespread and the consequences of heterogeneity should be no less a concern for psychologists. Second, unaccounted for heterogeneity is an indication of incomplete theory, since it suggests that a theory is unable to predict all contextual factors of importance to its claims. While heterogeneity reveals that a theory has flaws, by itself it does not damn a theory, and can be seen as an opportunity for theoretical advancement [ref: Tacket et al]. Third, heterogeneity can create controversy in the interpretation of replication results. The lack of a clear line for the degree of heterogeneity (susceptibility to contextual factors) that would make an effect uninteresting often results in heated discussions between original and replication authors. In response to this common area of disagreement [ref Simons et al] recently proposed that authors include a statement to explicitly delineate the level of heterogeneity that would make an effect lose its value. Fourth, meta-analytic techniques that attempt to correct for publication bias in their estimate tend to fail in the presence of heterogeneity [refs: Inzlicht; McShane]. This is problematic considering widespread publication bias [refs] and its impact on meta-analytic estimates. Finally, heterogeneity affects the interpretation of meta-analytic estimates as either _the_ true effect size (under homogeneity) or the average of the true effect sizes (under heterogeneity).

It is a common belief that heterogeneity is the norm in psychology. For example, several authors [ref: Tacket et al, McShane et al] argue that if we were to expect heterogeneity to be absent or minimal anywhere, it would be in pre-registered multi-lab projects with a common protocol, and that heterogeneity has been reported even under such circumstances [e.g., ref: Many Labs 1] is an indication of widespread heterogeneity in psychology. In addition, recent large scale reviews of meta-analyses in psychology [ref: Stanley; van Erp et al] report that much variability in effect sizes can be attributed to heterogeneity, with median estimates of 74% and 70.62% respectively [ref: Stanley]. In comparison, the median heterogeneity estimate in medicine is considerably lower: 21% amongst Cochrane meta-analyses [ref: Ioannidis  "uncertainty in heterogeneity"].

Heterogeneity is often measured by the I2 index [ref: Higgins & Higgins], which allows comparison of estimates across meta-analyses and has an intuitive interpretation. It can be interpreted as the percentage of variability in effect sizes that is due to heterogeneity amongst the true effects (that is, dependency on contextual factors) rather than sampling variance, and ranges from 0-100%. The I2 index is a transformation of the Q-test [ref: Cochran] by taking $[Q - (k - 1)] / Q \times 100$%, where k is the number of studies included in the meta-analysis and a negative I2 is set to zero.

Heterogeneity (I2) estimates typically contain considerable uncertainty. Both the Q-test and I2 have low power [ref: Jackson; Huedo-Medina]. This complicates the discussion of heterogeneity, because while I2 always provides an estimate of heterogeneity, this estimate is often surrounded by wide confidence intervals [ref: Ioannidis  "uncertainty in heterogeneity"]. For example, Ioannidis reports that in a large set of Cochrane meta-analyses, all meta-analyses with I2 point estimates of 0% had upper 95% confidence intervals that exceeded 33%. In addition, under homogeneity I2 has a central chisquare distribution [ref: von Hippel], a distribution that is right-skewed with more than 40% of observations falling above the expected value in almost all cases (for all k > 4). In other words, even in the absence of heterogeneity, a meta-analysis of 5 or more studies will have an I2 point estimate above zero in more than 40% of cases. Point estimates (expected values) of heterogeneity may thus be congruent with a wide range of true heterogeneity levels. Despite exhortations to the contrary [ref: Ioannidis "uncertainty in heterogeneity"] , it remains common to omit confidence intervals in the reporting of I2 [e.g., ML1 and ML3].

Effect size influences heterogeneity. Intuitively, it makes sense to think that if there is no meta-analytic effect there is nothing to moderate (i.e., no heterogeneity). However, a null or near null effect size estimate may arise from failure to consider contextual factors ('hidden moderators', [ref: van Bavel]) and does not by itself imply the absence of heterogeneity. This is especially true when the uncertainty in I2 estimates (see above) is taken into account. A large effect size on the other hand, can be expected to lead to more heterogeneity. To illustrate, consider a meta-analysis of say, the correlation between neuroticism and procrastination. Each included study would need to measure the two variables somehow, possibly the same way across studies. However, because of individual differences, measurement is not equally exact for everyone, and between studies the measurement reliabilities will differ either due to sampling variance (that the sample happens to be more or less homogenous) or to differences in contextual factors (e.g., sampling population, method of measurement). This means that even if the underlying effect size is the same, the observed correlation between the two variables will differ between studies. If measurement reliabilities stay the same, these differences will increase with the underlying effect, resulting in more variability being ascribed to heterogeneity. More formally, an observed correlation $r_{xy}$ is equivalent to the true correlation $\rho_{xy}$ multiplied by the square root of the measurement reliabilities for X ($R_{xx'}$) and Y ($R_{yy'}$): $r_{xy} = \rho_{xy} \times \sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$. As such, if $\sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$ stays constant while $\rho_{xy}$ increases the observed differences between studies will increase (see Table 2).

@@

```
We examine heterogeneity across 10 pre-registered multi-lab replication projects in psychology (Table 2). In light of the uncertainty surrounding I2-estimates we explore what conclusions can be drawn about heterogeneity in psychology from these projects, as well as the relationship between effect size and heterogeneity. These 10 projects examined a total of 37 primary outcome variables and arguably represent the best, least biased, meta-analytic data currently available in psychology. All included studies vary on a minimal number of contextual factors [as recognized by e.g., Tacket et al]: only sample population and settings. Nonetheless, contrary to the argument of [Tacket et al], if an effect is sensitive to changes in these factors we may still expect to find some heterogeneity. On the other hand, heterogeneity estimates will only inform us of susceptibility to changes in these two factors, with no information on susceptibility to changes in other contextual factors.

@@

summary.table %>% knitr::kable("latex", booktabs = T, digits = 0) %>% 
  kable_styling() %>% 
  column_spec(2, width = "25em") %>% #specify width of paper column
  footnote(threeparttable = TRUE,
           general = "For studies with several effects the number of participants is the average across effects, rounded to the closest whole number. Participant numbers are those used for primary analyses by original authors (i.e., after exclusions). RS = Replication Study, k = no. primary studies, ML = Many Labs, RRR = Registered Replication Report. (OSF.io/XXXX)")
```

@@

All code and data for this project are available on the Open Science Framework (OSF) at osf.io/XXXX. We refer to relevant files on OSF using brackets and links in the sections below, that is (osf.io/XXXX/). We ran all analyses using R version X [ref: R citation].

We downloaded and collated summary data from 10 pre-registered multi-lab replication projects in psychology (Table 2). Data from all 10 projects were publicly available on the Open Science Framework (osf.io) and downloaded between 2018/02/01 and 2018/03/31. Although some projects (e.g. RRR4) reported results from several outcome variables, we only included primary outcome variables as stated in accompanying publications, resulting in a total of 37 effects. For each effect we extracted (osf.io/XXXX/) summary data (e.g., means and standard deviations) at the level of the lab as specified by the original authors for their primary analysis (i.e., typically after exclusions). We extracted information on the country of each lab, whether participants were physically in the lab for the study, total number of participants per lab, type of effect size, and additional information related to each effect (see codebook; osf.io/XXXX/). Extracted data were in a variety of formats: Excel (Many labs 1, RRR1 & RRR2), CSV (Many labs 3, RRR3, RRR4, RRR5, RRR6) and as PDF tables (RRR7, RRR8). In two cases (RRR5 and RRR6) it was necessary to download the raw data to extract summary data. Although a particular lab may have participated in several projects, the lab indicator was typically not the same across projects. Even so, we kept the original lab indicators to facilitate comparing observations in our dataset with the original datasets. Finally, we collated he summary data for all effects into one dataset for analysis (osf.io/XXXX/).

To examine heterogeneity across effects, we computed meta-analytic estimates for all 37 effects in our dataset (Table 3). We ran all analyses as specified by the replication authors (osf.io/XXXX/). All effects were estimated with random-effects models and the Restricted Maximum Likelihood (REML) estimator using the R-package metafor [ref], though with a variety of outcome variables: correlations, standardized mean differences, raw mean differences, odds ratios and risk differences. ML1 transformed odd ratios into standardized mean differences when meta-analyzing under the assumption that responses followed logistic distributions [ref: Sanchez-Meca; metafor]. Two projects (RRR5 and RRR7) used the Knapp and Hartung adjustment of the standard errors [ref: Knapp & Hartung] and ML3 correlations were corrected for bias [ref: Hedges 1989; metafor]. After estimating effects, Ml3 transformed correlations into eta-squared for reporting, which we did not. For each effect we estimated I2 and its 95% confidence interval. Confidence intervals were calculated in metafor which uses the Q-profile method [ref Jackson 2014].

To examine the relationship between effect size and I2 we converted all effect sizes to a common metric (osf.io/XXXX/). Since I2 is logically and mathematically bounded at zero, we also correlated effect sizes with the closely related heterogeneity estimate H2, which is calculated as $H2 = Q / (k - 1)$. We first converted all effects into correlations (r) and used the R-package metafor [ref] to estimate I2 and meta-analytic effects expressed as Fisher's z. In doing so, we fit random-effects models with metafor's default REML estimator. Fisher's z is a normalizing correlation transformation that ranges from negative to positive infinity, though except for extreme correlations it stays close to the -1 to 1 range [ref]. For mean differences we calculated the pooled standard deviation, [ref Borenstein in Cooper & Hedges, 2009, p. 226), standardized the effect size and converted it to a correlation (p. 234) with a correction factor for unequal sample sizes (p. 234). In one case (RRR8), we first had to convert reported standard errors into pooled standard deviation (p.224). For risk differences and odds ratios we first added 1/2 to a cell if it was empty to avoid dividing by zero, next calculated the logarithmic odds ratio (p. 266), converted this to Cohen's d (p. 232) and then finally to a correlation. All formulas used are presented in Appendix A.

@@

##Table 3.
```{r heterogeneity-table}
#library(kableExtra)

het %>% knitr::kable("latex", align = "c", booktabs = TRUE,digits = 2, col.names = c("RS", "Effect", "k", "Effect type", "Estimate", "I2(%)", "ci.lb", "ci.ub")) %>% 
  kable_styling("scale_down") %>%
  footnote(threeparttable = TRUE, escape = FALSE,
           general = " Effects were estimated in metafor using REML. The following effects are odds ratios transformed into standardized mean differences: 'Allowed vs. forbidden', 'Gain vs. loss framing', 'Norm of reciprocity', 'Low vs. high category scales'. RS = Replication Study, k = no. primary studies, Estimate = Point estimates of effect sizes, ci.lb = lower bound of I2 95\\\\% confidence interval, ci.ub = upper bound of I2 95\\\\% confidence interval. SMD = Standardized Mean difference (Hedge's g), MD = Mean Difference, RD = Risk Difference, r = correlation. (OSF.io/XXXX)")

@@

Few effects show evidence of heterogeneity. The lower bound I2 confidence interval excludes zero for only 7/37 effects. Heterogeneity in a small number of effects is supported by the distribution of I2 point estimates. In a complete absence of heterogeneity, we would expect approximately 46% of point estimates across all effects to be non-zero, based on the chisquare distribution and average k across projects. That is, for about 17/37 effects. Table 3 shows that 23/37 effects have non-zero point estimates, indicating a small presence of heterogeneity.

Larger effect sizes appear to lead to higher heterogeneity estimates. In Table 3, the four effects with highest I2 estimates also have the largest effect sizes. As expected, heterogeneity tends to increase with (absolute) effect size, with effect size sufficient to explain 61% of variation in I2 estimates according to a simple regression model (Figure 1). The closely related, but unbounded, heterogeneity estimate H2 provides a similar result (R2 = .53).

@@

There is little evidence of heterogeneity amongst current pre-registered multi-lab replication projects. Across 10 projects, only 7/37 (19%) of primary outcomes showed evidence for heterogeneity. Moreover, while each replication of an effect only varied on two contextual factors (sampling population and settings), the effects that showed evidence for heterogeneity were primarily effects that might have been expected to be sensitive to the changes in sampling population. Save two effects (Anchoring - Everest and IAT correlation math), all other effects that demonstrated heterogeneity were related to the US. They either asked questions about the US (anchoring effects), persons related to the US (Quote attribution) or issues where the US is well-known to be exceptional (i.e., free speech; allowed vs. forbidden). Considering all seven effects belonged to the ML1 replication project, which collected data from labs in 10 different countries, these effects should have been expected to exhibit heterogeneity. [note to Jelte and Marcel: I am not sure this argumentation holds up, since ML1 did test US vs. non-US as a moderator and found very small effect sizes]

Complicating the discussion of heterogeneity is that there amongst the studied effects was a clear association between effect size and heterogeneity. There are thus both good theoretical (see introduction) and empirical reasons to expect larger effect sizes to exhibit comparatively more heterogeneity when using observed effect sizes in a meta-analysis. Since not all measurements (e.g., behavioral) admit adjusting for reliability as easily as questionnaires, this can be difficult to control for.

There are some limits to the generalizability of claims based on the data in our study. For one, the included effects are neither a representative nor random sample of effects in psychology and as such do not support making strong claims about average heterogeneity levels in psychology. In respect to this point we look forward to the forthcoming Many Labs 2 project which should provide considerable new data on the prevalence of heterogeneity in psychology. In addition, while both the I2 confidence intervals and distribution of point estimates support a small number of heterogenous effects amongst those studied, most confidence intervals are congruent with anything from low to high heterogeneity as defined by [Higgins et al]. As always, absence of evidence is not evidence of absence. Finally, since the included studies only vary on two contextual factors, the data tell us nothing about the included effects susceptibility to changes in the other two factors (treatment and measurement variables).

Evidence does not appear to support the belief that heterogeneity is the norm in psychology. Large-scale reviews of meta-analyses in psychology suggest considerable heterogeneity across the field of psychology [ref: Stanley; van Erp et al]. However, in our experience psychologist tend to include widely varying studies in their meta-analyses, sometimes varying on all four contextual factors (or more). This is perhaps unsurprising considering the lack of delineating theory in psychology [ref: Simons et al]. In support of this notion, [Ioannidis; uncertainty in heterogeneity] reports that, even after excluding meta-analyses of 3 or fewer studies, the median number of studies included in Cochrane meta-analyses is only 7, whereas in the dataset of [ref: van Erp et al] containing 747 meta-analyses in psychology, the median number of studies (effects) per meta-analysis is 12.]. Our examination of heterogeneity in pre-registered multi-lab replication projects, arguably the least biased meta-analytic data currently available in psychology, indicates that there is little evidence to support the notion that heterogeneity is widespread in psychology.

##References
